---
title: "5702 Project"
author: "Xinyuan He"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##loading in data from urls
```{r}
#install.packages("openxlsx")
library(openxlsx)

url = c("https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2019-03.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/03/LPR_data-2019-02.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/01/LPR_data-2019-01.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/01/LPR_data-2018-12.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-10.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-09.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-08.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-07.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-06.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-05.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-04.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-03.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-02.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2018-01.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-12.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-11.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-10.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-09.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-08.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-07.xlsx",
        "https://www.glassdoor.com/research/app/uploads/sites/2/2019/12/LPR_data-2017-06.xlsx")


#2018-11 can not be opened so not included in the url vector and not taken into account

i = 1
df_total = data.frame()
for(item in 1:length(url)){
  name = paste("df", i, sep = "")
  #print(name)
  a = read.xlsx(url[i])
  assign(name, a)
  df_total <- rbind(df_total,a)
  i = i+1
}
df_total
write.csv(df_total, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/df_total.csv")

#checking whether df_total contains all data from original datasets
nrow(df_total) == (nrow(df1) + nrow(df2)+ nrow(df3)+ nrow(df4)+ nrow(df5)+ nrow(df6)+ nrow(df7)+ nrow(df8)+ nrow(df9)+ nrow(df10)+ nrow(df11)+ nrow(df12)+ nrow(df13)+ nrow(df14)+ nrow(df15)+ nrow(df16)+ nrow(df17)+ nrow(df18)+ nrow(df19)+ + nrow(df20) + nrow(df21))
```

##loading in 2019 10 data from local csv
```{r}
#read in 2019.10 data for timeseries data
df_1910 = read.xlsx("/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/raw/JMR_data-2019-10.xlsx")
df_1910
```

##Timeseries 
national timeseries job opening is weekly data while national timeseries median pay is monthly data
city timeseries job opening is monthly data while city timeseries median pay is weekly data 
no much missing value in general
```{r}
#National + Timeseries, only have Job Openings under Dimension, no meidan pay data. weekly base
na_ts_opening = subset(df_1910, Metro == "National" & Dimension.Type == "Timeseries")
na_ts_opening
#US +Timeseries only timeseries median base pay, no job opening. monthly base 
na_ts_pay = subset(df_1910, Metro == "U.S."& Dimension.Type == "Timeseries")
na_ts_pay
#not able to combine na_ts and us_ts_pay since one weekly base and one monthly base


city_ts = subset(df_1910, Metro != "U.S." & Metro != "National" & Dimension.Type == "Timeseries")
drops <- c("Dimension.Type","YoY")
city_ts[drops]<- list(NULL)
city_ts

library(tidyr)
city_ts1 = city_ts %>% 
  spread(Measure, Value)
city_ts1
#job opening is monthly scale and pay is weekly scale 

#saving df as csv
write.csv(na_ts_opening, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/na_ts_opening.csv")
write.csv(na_ts_pay,"/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/na_ts_pay.csv")
write.csv(city_ts1,"/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/city_ts.csv")
```

##Job Title
Have only median pay data but no job opening data at all 
```{r}
#national job title, only pay, no job opening data
na_jobtitle = subset(df_total, Metro == "National" & Dimension.Type == "Job Title")
#names(na_jobtitle)
na_jobtitle
#na_jobtitle$Measure <- as.factor(na_jobtitle$Measure)
#levels(na_jobtitle$Measure)


#city job title, only pay, no job opening data
city_jobtitle = subset(df_total, Metro != "National" & Dimension.Type == "Job Title")
city_jobtitle
#city_jobtitle$Measure <- as.factor(city_jobtitle$Measure)
#levels(city_jobtitle$Measure)
#city_jobtitle$Metro <- as.factor(city_jobtitle$Metro)
#levels(city_jobtitle$Metro)

write.csv(na_jobtitle, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/na_jobtitle.csv")
write.csv(city_jobtitle, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/city_jobtitle.csv")

```

##Company Size
Have both median pay and job opening data 
problem is job opening data for earlier year are missing
```{r}
#national company size, with both pay and job opening
na_size = subset(df_total, Metro == "National" & Dimension.Type == "Company Size")
drops <- c("Dimension.Type","YoY")
na_size[drops]<- list(NULL)
na_size1 = na_size %>% 
  spread(Measure, Value)
na_size1

#city company size with both pay and job opening
city_size = subset(df_total, Metro != "National" & Dimension.Type == "Company Size")
drops <- c("Dimension.Type","YoY")
city_size[drops]<- list(NULL)
city_size1 = city_size %>% 
  spread(Measure, Value)
city_size1
#median pay data is pretty much complete while job opening missing lot of early data

write.csv(na_size1, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/na_size1.csv")
write.csv(city_size1, "/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/city_size1.csv")
```

##Industry
Have both median pay and job opening data 
problem is job opening data for earlier year are missing
```{r}
#national industry, with both pay and job opening
na_industry = subset(df_total, Metro == "National" & Dimension.Type == "Industry")
drops <- c("Dimension.Type","YoY")
na_industry[drops]<- list(NULL)
na_industry1 = na_industry %>% 
  spread(Measure, Value)
na_industry1
#median pay data is pretty much complete while job opening missing lot of early data


#city industry with both pay and job opening 
city_industry = subset(df_total, Metro != "National" & Dimension.Type == "Industry")
drops <- c("Dimension.Type","YoY")
city_industry[drops]<- list(NULL)
city_industry1 = city_industry %>% 
  spread(Measure, Value)
city_industry1
#median pay data is pretty much complete while job opening missing lot of early data


write.csv(na_industry1,"/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/na_industry1.csv")
write.csv(city_industry1,"/Users/xinyuanhe/Desktop/career-opportunities-analysis/data/clean/city_industry1.csv")
```